# Module 4 - Assignment 1
## Kelly Quesnel
### Missing Data Assignment Quiz

#### Libraries

```{r}
library(tidyverse)
library(VIM)
library(mice)
```

#### Read in data

```{r}
grades <- read_csv("class-grades.csv")
```

#### Task 1

Q: For which variables is there missing data?
A: Tutorial, Midterm, TakeHome, Final

Code:
```{r}
summary(grades)
```

#### Task 2

Q: Use the VIM package to visualize missingness. Are there any students that are missing multiple pieces of data?
A: Yes

Code:
```{r}
vim_plot <-
  aggr(grades, numbers = TRUE, prop = c(TRUE,FALSE),cex.axis=0.7)
```

#### Task 3

Q: Use row-wise deletion of missing values to create a new data frame. How many rows remain in this data frame?
A: 89

Code:
```{r}
grades_rowdel <- grades %>%
  drop_na()

nrow(grades_rowdel)
```

#### Task 4

Q: Use column-wise deletion of missing values to create a new data frame (from the original data frame not from the data frame created in Task 3). How many columns remain in this data frame?
A: 2

Code:
```{r}
grades_coldel <- grades %>%
  select(Prefix, Assignment)

ncol(grades_coldel)
```

#### Task 5

Q: Which approach (Task 3 or Task 4) seems preferable for this dataset? Briefly discuss your answer.
A: Task 3 (row-wise deletion) is a better choice for this data set. First of all, when you do column-wise deletion, it removes almost 2/3 of the data, whereas with the row-wise deletion, very few observations are deleted. Most importantly, column-wise deletion here actually removes the entire response variable (Final), which prevents us from doing any analysis.

#### Task 6

Q: Use the “mice” package to impute the missing data in this dataset. Set “m = 5” and “method=”pmm"" in the imputation. You MUST use a random number seed of 123 for this imputation. What is the mean value for the “Final” variable after imputation?
A: 68.33

Code: 
```{r}
set.seed(123)
imp_grades <-
  mice(grades, m=5, method='pmm', printFlag=FALSE)

summary(imp_grades)

grades_complete <-
  complete(imp_grades)

summary(grades_complete)
```











