# Module 3 - Assignment 1
## Kelly Quesnel
### Model Validation Assignment Quiz

Libraries
```{r}
library(tidyverse)
library(lubridate)
library(tidymodels)
```

Read in data
```{r}
bike <- read_csv("bike_cleaned-4.csv")
```
Clean data
```{r}
bike <- bike %>%
  mutate(dteday = mdy(dteday))

bike <- bike %>%
  mutate_if(is.character,as_factor)

bike <- bike %>%
  mutate(hr = as_factor(hr))
```

Split the data into training and testing sets. Your training set should have 70% of the data.
Use a random number (set.seed) of 1234. Your split should be stratified by the “count” variable.
```{r}
set.seed(1234)
bike_split <- initial_split(bike, prop = 0.7, strata = count)
bike_train <- training(bike_split)
bike_test <- testing(bike_split)

#Q1: How many rows of data are in the training set?
#A1: 12163

#Q2: Stratifying the split by the “count” variable serves what purpose?
#A2: Stratifying by “count” ensures that “count” is similarly represented in both the training and testing sets.
```

Build a linear regression model (using the training set) to predict “count” using the variables “season”, “mnth”, “hr”, “holiday”, and “weekday”, “temp”, and “weathersit”
```{r}
bike_recipe <-
  recipe(count ~ season + mnth + hr + holiday + weekday + temp + weathersit, bike_train)

bike_model <-
  linear_reg() %>%
  set_engine("lm")

bike_wflow <-
  workflow() %>%
  add_model(bike_model) %>%
  add_recipe(bike_recipe)

bike_fit <-
  fit(bike_wflow, bike_train)
```

```{r}
summary(bike_fit$fit$fit$fit)

#Q3: What is the adjusted R-squared value (to four digits) of the resulting model?
#A3: 0.6209
```

Use the predict functions to make predictions (using your model from Question 3) on the
training set. Hint: Be sure to store the predictions in an object, perhaps named “predict_train”
or similar. Develop a histogram of the predictions (Hint: The predictions are likely stored in a variable
called “.pred” in your predictions object).
```{r}
predict_train <- predict(bike_fit, new_data=bike_train)
ggplot(predict_train, aes(x=.pred)) +
  geom_histogram()
```
Select the statements below that are likely true about the distribution of predictions?
A. The maximum number of rides predicted for an hour is around 600 
B. The average number of rides predicted per hour is around 450
C. Some predictions for the number of rides in an hour are negative
D. None of these statements are true

A and C

Determine the performance of your model on the testing set.
```{r}
bike_fit %>% predict(bike_test) %>% bind_cols(bike_test) %>% metrics(truth = count, estimate = .pred)

#Q5: What is the R-squared value (to four decimal places) of your model on the testing set?
#A5: 0.6271
```

